---
title: "Rise Simulation Main"
author: "Arthur Hughes"
date: "2024-11-26"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Setup
```{r}
library(ggplot2)
library(dplyr)
library(quantreg)
library(Rsurrogate)
library(SurrogateRank)
library(MASS)
library(reshape2)
library(gridExtra)
library(cowplot)
library(grid)
library(pbmcapply)
library(tidyr)
library(RColorBrewer)
set.seed(23102024)
```

# Data Generation Function
Function to generate data. 
```{r}
gen.data <- function(n1, n0, p, prop_valid, valid_sigma, corr, mode = "simple") {
  if (mode == "simple"){
    
    # calculate the number of valid and invalid surrogates
    p_valid = prop_valid * p 
    p_invalid = (1 - prop_valid) * p
    
    # generate primary responses from multivariate normal distributions
    y1 <- rnorm(n1, y1_mean, y1_sd) # treated response 
    y0 <- rnorm(n0, y0_mean, y0_sd) # untreated repsonse
    
    # generate candidate surrogates 
    mm <- runif(p_invalid, min = 0.5, max = 2.5) # generate random invalid surrogate means from uniform distribution
    ss <- runif(p_invalid, min = 0.5, max = 2)  # generate random invalid surrogate sds from uniform distribution
    Sigma_invalid = matrix(corr, nrow = p_invalid, ncol= p_invalid) # Invalid surrogate covariance matrix
    diag(Sigma_invalid) = ss 
    
    if (prop_valid != 0){ # if valid surrogates required
      
      # valid surrogate covariance matrix
      Sigma_valid = matrix(corr*valid_sigma, nrow = p_valid, ncol= p_valid) 
      diag(Sigma_valid) = rep(valid_sigma, p_valid)
      
      # valid surrogates in treated group by perturbing primary response
      s1.valid <- matrix(y1, nrow = n1, ncol = p_valid, byrow = TRUE) + 
        mvrnorm(n = n1, mu = rep(0, p_valid), Sigma = Sigma_valid)
      
      # valid surrogates in untreated group by perturbing primary response
      s0.valid <- matrix(y0, nrow = n0, ncol = p_valid, byrow = TRUE) + 
        mvrnorm(n = n0, mu = rep(0, p_valid), Sigma = Sigma_valid)
      
      # invalid surrogates
      if (prop_valid != 1){ # if invalid surrogates required
        s1.invalid <- mvrnorm(n = n1, mu = mm, Sigma = Sigma_invalid)
        s0.invalid <- mvrnorm(n = n0, mu = mm, Sigma = Sigma_invalid)
        # bind candidates together 
        s1 <- cbind(s1.valid, s1.invalid)
        s0 <- cbind(s0.valid, s0.invalid)
      } else {
        s1 = s1.valid 
        s0 = s0.valid 
      }
      
    } else { # if no valid surrogates required 
      # invalid surrogates
      s1 <- mvrnorm(n = n1, mu = mm, Sigma = Sigma_invalid)
      s0 <- mvrnorm(n = n0, mu = mm, Sigma = Sigma_invalid)
    }
    # store hypothesis truths
    hyp <- c(rep("null false", p_valid), rep("null true", p_invalid))
    
    # return generated data as a list
    return(list(y1 = y1, y0 = y0, s1 = s1, s0 = s0, hyp = hyp))
  } else if (mode == "complex") {
    # calculate the number of valid and invalid surrogates
    p_valid = prop_valid * p 
    p_invalid = (1 - prop_valid) * p
    
    # generate primary responses from multivariate normal distributions
    y1 <- rnorm(n1, y1_mean, y1_sd) # treated response 
    y0 <- rnorm(n0, y0_mean, y0_sd) # untreated repsonse
    
    # generate candidate surrogates 
    lambda = runif(p_invalid, min = 0.5, max = 2.5)
    if (prop_valid != 0){ #if valid surrogates required
      # valid surrogate covariance matrix
      Sigma_valid = matrix(corr*valid_sigma, nrow = p_valid, ncol= p_valid) 
      diag(Sigma_valid) = rep(valid_sigma, p_valid)
      s1.valid = matrix(y1^3, nrow = n1, ncol = p_valid, byrow = TRUE) + 
        mvrnorm(n = n1, mu = rep(0, p_valid), Sigma = Sigma_valid)
      
      s0.valid = matrix(y0^3, nrow = n0, ncol = p_valid, byrow = TRUE) + 
        mvrnorm(n = n0, mu = rep(0, p_valid), Sigma = Sigma_valid)
      
      if (prop_valid != 1){ #if invalid surrogates required
        s0.invalid = sapply(lambda, function(rate) rexp(n0, rate))
        s1.invalid = sapply(lambda, function(rate) rexp(n1, rate))
        s1 <- cbind(s1.valid, s1.invalid)
        s0 <- cbind(s0.valid, s0.invalid)
      } else {
        s1 = s1.valid 
        s0 = s0.valid  
      }
    } else {
      s0 = sapply(lambda, function(rate) rexp(n0, rate))
      s1 = sapply(lambda, function(rate) rexp(n1, rate))
    }
    hyp <- c(rep("null false", p_valid), rep("null true", p_invalid))
    return(list(y1 = y1, y0 = y0, s1 = s1, s0 = s0, hyp = hyp))
  }
}
```

# Truth function
Function to calculate the true test statistics using asymptotic properties of U statistics. 
```{r}
calc.truth <- function(p, prop_valid, valid_sigma, corr, mode = "simple") {
  # generate a dataset with a large sample size 
  dd <- gen.data(n1 = 10000, n0 = 10000, p, prop_valid, valid_sigma, corr = corr, mode = mode)
  
  # find the asymptotic U statistic for Y
  uy <- (10000 * 10000)^(-1) * wilcox.test(dd$y1, dd$y0)$statistic
  
  # find the asymptotic U statistic for each surrogate
  us <- numeric(length(which(dd[["hyp"]] == "null false")))
  for (j in which(dd[["hyp"]] == "null false")) {
    us[j] <- (10000 * 10000)^(-1) * wilcox.test(dd$s1[, j], dd$s0[, j])$statistic
  }
  
  # take the truth as the mean difference
  delta_mean <- mean(uy - us)
  delta_sd <- sd(uy - us)
  return(list(uy_true = uy, 
              us_true = us, 
              delta_true = delta_mean, 
              delta_true_sd = delta_sd))
}
```

# Evaluation function
Function which repeatedly simulates data and calculates performance metrics (FPR, FDP, TPR).
```{r}
# Function to simulate results (performance metrics)
simulate_results <- function(n1, n0, p, prop_valid, n_sim, valid_sigma, corr, mode = "simple") {
  p_unadjusted <- matrix(0, nrow = n_sim, ncol = p) # store unadjusted p_values for each candidate and for each simulation
  fpr_unadjusted <- numeric(n_sim) # store unadjusted false positive rates per simulation
  fdr_unadjusted <- numeric(n_sim) # store unadjusted proportion of false positives per simulation
  tpr_unadjusted <- numeric(n_sim) # store unadjusted true positive rates (empirical power) per simulation
  ppv_unadjusted <- numeric(n_sim) # store unadjusted positive predictive values per simulation
  
  p_bonf <- matrix(0, nrow = n_sim, ncol = p) # store bonferroni corrected p-values
  fpr_bonf <- numeric(n_sim)
  fdr_bonf <- numeric(n_sim)
  tpr_bonf <- numeric(n_sim)
  ppv_bonf <- numeric(n_sim)
  
  p_bh <- matrix(0, nrow = n_sim, ncol = p) # store benjamini-hochberg corrected p-values
  fpr_bh <- numeric(n_sim)
  fdr_bh <- numeric(n_sim)
  tpr_bh <- numeric(n_sim)
  ppv_bh <- numeric(n_sim)
  
  p_by <- matrix(0, nrow = n_sim, ncol = p) # store benjamini-yekutieli corrected p-values
  fpr_by <- numeric(n_sim)
  fdr_by <- numeric(n_sim)
  tpr_by <- numeric(n_sim)
  ppv_by <- numeric(n_sim)
  
  for (k in 1:n_sim) { # for loop to do many simulations
    # generate data
    data <- gen.data(n1, n0, p, prop_valid, valid_sigma, corr, mode = mode)
    
    # Estimate u_y
    u_y_estimated <- test.surrogate(
      yone = data$y1, yzero = data$y0, sone = data$y1, szero = data$y0, epsilon = 0.1
    )$u.y
    # u_y_true = truth$uy_true
    # choose epsilon as maximum of estimated value of UY - 0.5 and 0
    eps <- max(0, u_y_estimated - 0.5)
    
    for (j in 1:p) {  # for each candidate surrogate 
      ss.test <- test.surrogate(
        yone = data$y1, 
        yzero = data$y0, 
        sone = data$s1[, j], # jth candidate surrogate 
        szero = data$s0[, j], 
        epsilon = eps)
      # compute unadjusted p-value for jth candidate surrogate
      p_unadjusted[k, j] <- pnorm(ss.test$delta.estimate, ss.test$epsilon.used, ss.test$sd.delta)
    }
    # correct p-values per-simulation with bonferroni, BH, and BY. 
    p_bonf[k, ] = p.adjust(p_unadjusted[k, ], method = "bonferroni")
    p_bh[k, ] = p.adjust(p_unadjusted[k, ], method = "BH")
    p_by[k, ] = p.adjust(p_unadjusted[k, ], method = "BY")
    
    
    p_values_unadjusted = p_unadjusted[k, ]
    # Classify test outcomes for uncorrected p-values
    TP_unadjusted <- sum(data[["hyp"]] == "null false" & p_values_unadjusted < alpha) # true positives
    FP_unadjusted <- sum(data[["hyp"]] == "null true" & p_values_unadjusted < alpha) # false positives
    TN_unadjusted <- sum(data[["hyp"]] == "null true" & p_values_unadjusted >= alpha) # true negatives
    FN_unadjusted <- sum(data[["hyp"]] == "null false" & p_values_unadjusted >= alpha) # false negatives
    
    # repeat for each correction method
    p_values_bonf = p_bonf[k, ]
    TP_bonf <- sum(data[["hyp"]] == "null false" & p_values_bonf < alpha)
    FP_bonf <- sum(data[["hyp"]] == "null true" & p_values_bonf < alpha)
    TN_bonf <- sum(data[["hyp"]] == "null true" & p_values_bonf >= alpha)
    FN_bonf <- sum(data[["hyp"]] == "null false" & p_values_bonf >= alpha)
    
    p_values_bh = p_bh[k, ]
    TP_bh <- sum(data[["hyp"]] == "null false" & p_values_bh < alpha)
    FP_bh <- sum(data[["hyp"]] == "null true" & p_values_bh < alpha)
    TN_bh <- sum(data[["hyp"]] == "null true" & p_values_bh >= alpha)
    FN_bh <- sum(data[["hyp"]] == "null false" & p_values_bh >= alpha)
    
    p_values_by = p_by[k, ]
    TP_by <- sum(data[["hyp"]] == "null false" & p_values_by < alpha)
    FP_by <- sum(data[["hyp"]] == "null true" & p_values_by < alpha)
    TN_by <- sum(data[["hyp"]] == "null true" & p_values_by >= alpha)
    FN_by <- sum(data[["hyp"]] == "null false" & p_values_by >= alpha)
    
    
    if (prop_valid == 0){ # if no true surrogates we only calculate the false positive rate
      fpr_unadjusted[k] <- FP_unadjusted / (FP_unadjusted + TN_unadjusted) # FPR = FP/(FP + TN)                 # Proportion of false positives
      fdr_unadjusted[k] <- 0  # set the other metrics as 0
      tpr_unadjusted[k] <- 0                 
      ppv_unadjusted[k] <- 0
    } else if (prop_valid != 0){  # if any true positives, calculate FPR, FDP, TPR, PPV
      fpr_unadjusted[k] <- FP_unadjusted / (FP_unadjusted + TN_unadjusted)  # False positive rate = FP/(FP + TN)   
      fdr_unadjusted[k] <- FP_unadjusted / (TP_unadjusted + FP_unadjusted)  # Proportion of False Positives = FP/(FP + TP) 
      tpr_unadjusted[k] <- TP_unadjusted / (TP_unadjusted + FN_unadjusted)  # True Positive Rate (Sensitivity) = TP/(TP + FN)
      ppv_unadjusted[k] <- TP_unadjusted / (TP_unadjusted + FP_unadjusted)  # Positive Predictive Value = TP/(TP + FP)
      
      # repeat for all the correction methods
      fpr_bonf[k] <- FP_bonf / (FP_bonf + TN_bonf)                 
      fdr_bonf[k] <- FP_bonf / (TP_bonf + FP_bonf)                  
      tpr_bonf[k] <- TP_bonf / (TP_bonf + FN_bonf)                
      ppv_bonf[k] <- TP_bonf / (TP_bonf + FP_bonf)                 
      
      fpr_bh[k] <- FP_bh / (FP_bh + TN_bh)                
      fdr_bh[k] <- FP_bh / (TP_bh + FP_bh)                 
      tpr_bh[k] <- TP_bh / (TP_bh + FN_bh)                  
      ppv_bh[k] <- TP_bh / (TP_bh + FP_bh)                
      
      fpr_by[k] <- FP_by / (FP_by + TN_by)                 
      fdr_by[k] <- FP_by / (TP_by + FP_by)                
      tpr_by[k] <- TP_by / (TP_by + FN_by)                 
      ppv_by[k] <- TP_by / (TP_by + FP_by)                 
    }
    # print progress within simulation runs
    # print(paste0("Sigma = ", valid_sigma, ", p = ", p, ", n = ", n, " : simulation ", k, " of ", n_sim))
  }
  
  # Where we have NA in our metrics due to division by 0 
  # (i.e. in particular, where we do not identify any positives)
  # replace with a 0 
  fpr_unadjusted[is.nan(fpr_unadjusted)] <- 0
  fdr_unadjusted[is.nan(fdr_unadjusted)] <- 0
  tpr_unadjusted[is.nan(tpr_unadjusted)] <- 0
  ppv_unadjusted[is.nan(ppv_unadjusted)] <- 0
  
  fpr_bonf[is.nan(fpr_bonf)] <- 0
  fdr_bonf[is.nan(fdr_bonf)] <- 0
  tpr_bonf[is.nan(tpr_bonf)] <- 0
  ppv_bonf[is.nan(ppv_bonf)] <- 0
  
  fpr_bh[is.nan(fpr_bh)] <- 0
  fdr_bh[is.nan(fdr_bh)] <- 0
  tpr_bh[is.nan(tpr_bh)] <- 0
  ppv_bh[is.nan(ppv_bh)] <- 0
  
  fpr_by[is.nan(fpr_by)] <- 0
  fdr_by[is.nan(fdr_by)] <- 0
  tpr_by[is.nan(tpr_by)] <- 0
  ppv_by[is.nan(ppv_by)] <- 0
  
  # calculate the mean metrics across simulations
  avg_fpr_unadjusted <- mean(fpr_unadjusted)
  avg_fdr_unadjusted <- mean(fdr_unadjusted)
  avg_tpr_unadjusted <- mean(tpr_unadjusted)
  avg_ppv_unadjusted <- mean(ppv_unadjusted)
  
  avg_fpr_bonf <- mean(fpr_bonf)
  avg_fdr_bonf <- mean(fdr_bonf)
  avg_tpr_bonf <- mean(tpr_bonf)
  avg_ppv_bonf <- mean(ppv_bonf)
  
  avg_fpr_bh <- mean(fpr_bh)
  avg_fdr_bh <- mean(fdr_bh)
  avg_tpr_bh <- mean(tpr_bh)
  avg_ppv_bh <- mean(ppv_bh)
  
  avg_fpr_by <- mean(fpr_by)
  avg_fdr_by <- mean(fdr_by)
  avg_tpr_by <- mean(tpr_by)
  avg_ppv_by <- mean(ppv_by)
  
  # store the metrics in lists per correction method
  metrics_unadjusted = list("avg_fpr" = avg_fpr_unadjusted,
                            "avg_fdr" = avg_fdr_unadjusted,
                            "avg_tpr" = avg_tpr_unadjusted,
                            "avg_ppv" = avg_ppv_unadjusted)
  
  metrics_bonf = list("avg_fpr" = avg_fpr_bonf,
                      "avg_fdr" = avg_fdr_bonf,
                      "avg_tpr" = avg_tpr_bonf,
                      "avg_ppv" = avg_ppv_bonf)
  
  metrics_bh = list("avg_fpr" = avg_fpr_bh,
                    "avg_fdr" = avg_fdr_bh,
                    "avg_tpr" = avg_tpr_bh,
                    "avg_ppv" = avg_ppv_bh)
  
  metrics_by = list("avg_fpr" = avg_fpr_by,
                    "avg_fdr" = avg_fdr_by,
                    "avg_tpr" = avg_tpr_by,
                    "avg_ppv" = avg_ppv_by)
  
  metrics = list("metrics_unadjusted" = metrics_unadjusted, 
                 "metrics_bonf" = metrics_bonf,
                 "metrics_bh" = metrics_bh,
                 "metrics_by" = metrics_by)
  
  # print progress of simulation number
  # print(paste0("p = ", p, ", n = ", n, " : simulation complete!"))
  
  p_values = list("p_unadjusted" = p_unadjusted, 
                  "p_bonf" = p_bonf,
                  "p_bh" = p_bh,
                  "p_by" = p_by)
    # output of function are the p-value lists and the performance metrics
    return(list("p_values" = p_values, "metrics"= metrics))
}
```
# Main 
## Figure 1
boxplots of FPR as a function of sample size in the uncorrelated setting.
```{r}
# Set parameters
alpha <- 0.05 # nominal significance level
beta <- 0.2 # desired power for univariate surrogate test = (1-beta)*100% 
s0_mean <- 0 # invalid surrogate mean value in untreated group
s0_sd <- 1 # invalid surrogate sd value in untreated group
s1_mean <- 0 # invalid surrogate mean value in treated group
s1_sd <- 2 # invalid surrogate sd value in treated group
y0_mean <- 0 # primary response mean value in untreated group
y0_sd <- 1 # primary response sd value in untreated group
y1_mean <- 3 # primary response mean value in treated group
y1_sd <- 1 # primary response sd value in treated group
p <- 500 # total number of predictors
prop_valid <- 0 # proportion of predictors which are valid surrogates
p_valid <- prop_valid * p # number of valid surrogates
p_invalid = (1 - prop_valid) * p # number of invalid surrogates
n_sim = 500 # number of simulations
corr = 0

n_grid = seq(10,100,10)

metrics_fig1 = matrix(0,nrow = length(n_grid)*n_sim, ncol = 2)

for (i in 1:length(n_grid)){
  for (j in 1:n_sim){
    res <- simulate_results(n1 = n_grid[i] / 2, 
                            n0 = n_grid[i] / 2, 
                            p = p, 
                            prop_valid = 0, 
                            n_sim = 1, 
                            valid_sigma,
                            corr = 0, 
                            mode = "simple")
    
    # p_values = append(p_values,list(res[["p_values"]][["p_unadjusted"]]))
    
    # Append results to data frame
    
    metrics_fig1[(i-1)*(n_sim) + j,1] = n_grid[i]
    metrics_fig1[(i-1)*(n_sim) + j,2] = res[["metrics"]][["metrics_unadjusted"]][["avg_fpr"]]
    print(paste0("Sample size : ", n_grid[i], ", simulation ", j, " of ", n_sim, " complete." ))
  }
}

metrics_fig1 = data.frame("n" = metrics_fig1[,1],
                          "False Positive Rate" = metrics_fig1[,2])

save(metrics_fig1, file = "./output/simulation/metrics_figure1.Rdata")
metrics_figure1 = get(load("./output/simulation/metrics_figure1.Rdata"))

fig1 = metrics_fig1 %>% ggplot(aes(x = as.factor(n), 
                                   y = FPR,
                                   fill = as.factor(n))) +
  geom_boxplot() +
  labs(title = "Observed FPR across different sample sizes",
       x = "Sample size",
       y = "False Positive Rate",
       fill = "Sample Size") + 
  ylim(0,0.15) +
  geom_hline(yintercept = 0.05, colour = "red", linetype = "dashed", linewidth = 1.1) +
  theme_minimal(base_size = 35) +
  theme(plot.title = element_blank(),
        axis.title = element_text(size = 35),
        plot.background = element_rect(fill = 'white', color = 'white'),
        legend.position = "none") +
  scale_fill_brewer(palette = "RdYlGn")

fig1

ggsave(plot = fig1,
       "./output/figures/figure1.eps", 
       units = "cm", 
       width = 40, 
       height = 25, dpi = 800)
```

## Figure 2
Violin plots of FPR as a function of correlation.
```{r}
# Set parameters
alpha <- 0.05 # nominal significance level
beta <- 0.2 # desired power for univariate surrogate test = (1-beta)*100% 
s0_mean <- 0 # invalid surrogate mean value in untreated group
s0_sd <- 1 # invalid surrogate sd value in untreated group
s1_mean <- 0 # invalid surrogate mean value in treated group
s1_sd <- 2 # invalid surrogate sd value in treated group
y0_mean <- 0 # primary response mean value in untreated group
y0_sd <- 1 # primary response sd value in untreated group
y1_mean <- 3 # primary response mean value in treated group
y1_sd <- 1 # primary response sd value in treated group
n <- 50 # total sample size
n0 <- n / 2 # treated sample size 
n1 <- n / 2 # untreated sample size 
p <- 500 # total number of predictors
prop_valid <- 0 # proportion of predictors which are valid surrogates
p_valid <- prop_valid * p # number of valid surrogates
p_invalid = (1 - prop_valid) * p # number of invalid surrogates
n_sim = 500 # number of simulations

corr_grid = seq(0,0.5,0.1)

metrics_fig2 = matrix(0,nrow = length(corr_grid)*n_sim, ncol = 2)

for (i in 1:length(corr_grid)){
  for (j in 1:n_sim){
      res <- simulate_results(n1 = n / 2, 
                              n0 = n / 2, 
                              p = p, 
                              prop_valid = 0, 
                              n_sim = 1, 
                              corr = corr_grid[i], 
                              mode = "simple")
      
      # p_values = append(p_values,list(res[["p_values"]][["p_unadjusted"]]))
      
      # Append results to data frame
      
      metrics_fig2[(i-1)*(n_sim) + j,1] = corr_grid[i]
      metrics_fig2[(i-1)*(n_sim) + j,2] = res[["metrics"]][["metrics_unadjusted"]][["avg_fpr"]]
      print(paste0("Correlation : ", corr_grid[i], ", simulation ", j, " of ", n_sim, " complete." ))
      }
}

metrics_fig2 = data.frame("correlation" = metrics_fig2[,1],
                          "False Positive Rate" = metrics_fig2[,2])

save(metrics_fig2, file = "./output/simulation/metrics_figure2.Rdata")

metrics_fig2 = get(load("./output/simulation/metrics_figure2.Rdata"))

fig2 = metrics_fig2 %>% ggplot(aes(x = as.factor(correlation), 
                                   y = FPR,
                                   fill = as.factor(correlation))) +
  geom_violin() +
  labs(title = "Observed FPR across different inter-predictor correlation levels",
       x = expression(sigma[corr]),
       y = "False Positive Rate",
       fill = "Correlation") + 
  geom_hline(yintercept = 0.05, colour = "red", linetype = "dashed", linewidth = 1.1) +
  theme_minimal(base_size = 35) +
  theme(plot.title = element_blank(),
        plot.background = element_rect(fill = 'white', color = 'white'),
        legend.position = "none",
        axis.title.x = element_text(size = 55)) +
  scale_fill_brewer(palette = "RdYlGn")

fig2

ggsave(plot = fig2,
       "./output/figures/figure2.eps", 
       units = "cm", 
       width = 40, 
       height = 25, dpi = 800)
```

## Figure 3
Empirical power and proportion of false positives as a function of average surrogate strength for three different sample sizes. Average surrogate strengths corresponding to (0.55,0.6,...,0.95) are found by trial-and-error.
```{r}
# round(mean(calc.truth(p, prop_valid, valid_sigma = 244, corr, mode = "simple")$us_true),2)
# round(mean(calc.truth(p, prop_valid, valid_sigma = 68, corr, mode = "simple")$us_true),2)
# round(mean(calc.truth(p, prop_valid, valid_sigma = 30, corr, mode = "simple")$us_true),2)
# round(mean(calc.truth(p, prop_valid, valid_sigma = 15, corr, mode = "simple")$us_true),2)
# round(mean(calc.truth(p, prop_valid, valid_sigma = 9, corr, mode = "simple")$us_true),2)
# round(mean(calc.truth(p, prop_valid, valid_sigma = 5.5, corr, mode = "simple")$us_true),2)
# round(mean(calc.truth(p, prop_valid, valid_sigma = 3, corr, mode = "simple")$us_true),2)
# round(mean(calc.truth(p, prop_valid, valid_sigma = 1.8, corr, mode = "simple")$us_true),2)
# round(mean(calc.truth(p, prop_valid, valid_sigma = 0.65, corr, mode = "simple")$us_true),2)
# round(mean(calc.truth(p, prop_valid, valid_sigma = 0.01, corr, mode = "simple")$us_true),2)

# Set up the values to simulate
alpha <- 0.05 # nominal significance level
beta <- 0.2 # desired power for univariate surrogate test = (1-beta)*100% 
s0_mean <- 0 # invalid surrogate mean value in untreated group
s0_sd <- 1 # invalid surrogate sd value in untreated group
s1_mean <- 0 # invalid surrogate mean value in treated group
s1_sd <- 2 # invalid surrogate sd value in treated group
y0_mean <- 0 # primary response mean value in untreated group
y0_sd <- 1 # primary response sd value in untreated group
y1_mean <- 3 # primary response mean value in treated group
y1_sd <- 1 # primary response sd value in treated group
p <- 500 # total number of predictors
sigma_grid = c(0.01,0.65,1.8,3,5.5,9,15,30,68,244) # values of sigma
n_grid <- c(30,50,100) # 3 sample sizes
n_sim = 500
prop_valid = 0.1
corr = 0

# Initialise an empty list for p-values and dataframe for metrics
p_values = list()
metrics_fig3 <- data.frame(
  n = integer(), 
  p = integer(), 
  sigma_S = numeric(),
  avg_us = numeric(),
  avg_fpr_unadjusted = numeric(),
  avg_fdr_unadjusted = numeric(),
  avg_tpr_unadjusted = numeric(),
  avg_ppv_unadjusted = numeric(),
  avg_fpr_bonf = numeric(),
  avg_fdr_bonf = numeric(),
  avg_tpr_bonf = numeric(),
  avg_ppv_bonf = numeric(),
  avg_fpr_bh = numeric(),
  avg_fdr_bh = numeric(),
  avg_tpr_bh = numeric(),
  avg_ppv_bh = numeric(),
  avg_fpr_by = numeric(),
  avg_fdr_by = numeric(),
  avg_tpr_by = numeric(),
  avg_ppv_by = numeric()
)

# Perform simulations over grid
for (sigma in sigma_grid){
  avg_us = round(mean(calc.truth(p, prop_valid, valid_sigma = sigma, corr, mode = "simple")$us_true),2)
  for (n in n_grid) {
      res <- simulate_results(n1 = n / 2, 
                              n0 = n / 2, 
                              p = p, 
                              prop_valid, 
                              n_sim = n_sim, 
                              valid_sigma = sigma,
                              corr = 0, 
                              mode = "simple")
      
      # p_values = append(p_values,list(res[["p_values"]][["p_unadjusted"]]))
      
      # Append results to data frame
      metrics_fig3 <- rbind(metrics_fig3, data.frame(
        n = n, p = p,
        sigma_S = sigma,
        avg_us = avg_us,
        avg_fpr_unadjusted = res[["metrics"]][["metrics_unadjusted"]][["avg_fpr"]], 
        avg_fdr_unadjusted = res[["metrics"]][["metrics_unadjusted"]][["avg_fdr"]], 
        avg_tpr_unadjusted = res[["metrics"]][["metrics_unadjusted"]][["avg_tpr"]],
        avg_ppv_unadjusted = res[["metrics"]][["metrics_unadjusted"]][["avg_ppv"]],
        avg_fpr_bonf = res[["metrics"]][["metrics_bonf"]][["avg_fpr"]], 
        avg_fdr_bonf = res[["metrics"]][["metrics_bonf"]][["avg_fdr"]], 
        avg_tpr_bonf = res[["metrics"]][["metrics_bonf"]][["avg_tpr"]],
        avg_ppv_bonf = res[["metrics"]][["metrics_bonf"]][["avg_ppv"]],
        avg_fpr_bh = res[["metrics"]][["metrics_bh"]][["avg_fpr"]], 
        avg_fdr_bh = res[["metrics"]][["metrics_bh"]][["avg_fdr"]], 
        avg_tpr_bh = res[["metrics"]][["metrics_bh"]][["avg_tpr"]],
        avg_ppv_bh = res[["metrics"]][["metrics_bh"]][["avg_ppv"]],
        avg_fpr_by = res[["metrics"]][["metrics_by"]][["avg_fpr"]], 
        avg_fdr_by = res[["metrics"]][["metrics_by"]][["avg_fdr"]], 
        avg_tpr_by = res[["metrics"]][["metrics_by"]][["avg_tpr"]],
        avg_ppv_by = res[["metrics"]][["metrics_by"]][["avg_ppv"]]
      ))
    print(n)
  }
}

save(metrics_fig3, file = "./output/simulation/metrics_figure3.Rdata")

metrics_fig3 = get(load("./output/simulation/metrics_figure3.Rdata"))

# FDR as a function of surrogate strength
FDR_plot_unadjusted = metrics_fig3 %>% ggplot(aes(x = avg_us, y = avg_fdr_unadjusted, color = as.factor(n), linetype = as.factor(n))) + 
  geom_line(size = 1.3) + 
  geom_point() +
  ylab("False Discovery Proportion") +
  xlab(expression(bar(U[S]))) +
  scale_color_brewer(palette = "Set1") + 
  guides(color = guide_legend(title = "Sample size")) +
  ggtitle("Proportion of False Positives as a function of average surrogate strength - unadjusted p-values") +
  theme_minimal(base_size = 24) +
  theme(plot.background = element_rect(fill = 'white', color = 'white'),
        plot.title = element_blank()) 

TPR_plot_unadjusted  = metrics_fig3 %>% ggplot(aes(x = avg_us, y = avg_tpr_unadjusted , color = as.factor(n), linetype = as.factor(n))) + 
  geom_line(size = 1.3) + 
  geom_point() +
  ylab("Empirical power") +
  xlab(expression(bar(U[S]))) +
  scale_color_brewer(palette = "Set1") + 
  guides(color = guide_legend(title = "Sample size")) +
  ggtitle("Empirical power as a function of average surrogate strength - unadjusted p-values") +
  theme_minimal(base_size = 24) +
  theme(plot.background = element_rect(fill = 'white', color = 'white'),
        plot.title = element_blank()) 

## FPR and TPR in the same grid 
TPR_patchwork = TPR_plot_unadjusted + 
  ylim(0,1) + 
  theme(legend.position = "none")

FDR_patchwork = FDR_plot_unadjusted + 
  ylim(0,1) + 
  theme(legend.position = "none")

# Extract the legend
legend <- get_legend(
  metrics_fig3 %>% 
    ggplot(aes(x = avg_us, 
               y = avg_fdr_unadjusted, 
               color = as.factor(n), 
               linetype = as.factor(n))) + 
    geom_line(size = 1.3) + # Adjust the line size in the plot
    geom_point() +
    ylab("False Discovery Proportion") +
    xlab(expression(bar(U[S]))) +
    scale_color_brewer(palette = "Set1") + 
    guides(
      color = guide_legend(title = "Sample size"),
      linetype = guide_legend(
        title = "Sample size",
        override.aes = list(size = 1) # Adjust line size in legend
      )
    ) +
    ggtitle("Proportion of False Positives as a function of average surrogate strength - unadjusted p-values") +
    theme_minimal(base_size = 20) +
    theme(
      legend.key.width = unit(2, "cm") # Increase legend key width
    )
)

figure3 = grid.arrange(
  arrangeGrob(TPR_patchwork, FDR_patchwork, legend, 
              ncol = 3, widths = c(1, 1, 0.3)),  # Adjust widths as needed
  top = textGrob("",
                 gp = gpar(fontsize = 20))  # Adjust fontsize and fontface
)

ggsave(plot = figure3,
       "./output/figures/figure3.eps", 
       units = "cm", 
       width = 35, 
       height = 18, dpi = 800)
```

## Figure 4
Evaluation stage : Vary the proportion of false positives in the combined marker and examine distribution of p-values.
```{r}
# Step 1 : construct a marker as a combination of true and false positives
# Keep true positives strong on average : U_Y = 0.9
source("./R/rise_validate.R")
source("./R/rise_screen.R")
alpha <- 0.05 # nominal significance level
beta <- 0.2 # desired power for univariate surrogate test = (1-beta)*100% 
s0_mean <- 0 # invalid surrogate mean value in untreated group
s0_sd <- 1 # invalid surrogate sd value in untreated group
s1_mean <- 0 # invalid surrogate mean value in treated group
s1_sd <- 2 # invalid surrogate sd value in treated group
y0_mean <- 0 # primary response mean value in untreated group
y0_sd <- 1 # primary response sd value in untreated group
y1_mean <- 3 # primary response mean value in treated group
y1_sd <- 1 # primary response sd value in treated group
p <- 20 # total number of predictors
corr = 0 # inter-predictor correlation
n1 = 25 # treated individuals
n0 = 25 # untreated
p = 20 # number of genes to combine for the signature
valid_sigma = 1.8 # this corresponds to avg U_Y = 0.9
prop_invalid_grid = seq(0,1,0.1) # proportions of invalid surrogates in gamma to test
n_sim = 500 # number of simulations

# evaluate true surrogate strength
# truth <- calc.truth(p = 500, prop_valid=1, valid_sigma, corr, mode = "simple")
# mean(truth$us_true)

p_evaluate_df = as.data.frame(matrix(0,nrow = n_sim, ncol = length(prop_invalid_grid)))
colnames(p_evaluate_df) = prop_invalid_grid
for (i in 1:length(prop_invalid_grid)){
  for (j in 1:n_sim){
    prop_invalid = prop_invalid_grid[i]
    prop_valid = 1 - prop_invalid 
    
    data = gen.data(n1, n0, p, prop_valid, valid_sigma, corr, mode = "simple")
    
    # get weights from screening function
    Y = c(data$y1,data$y0)
    A = c(rep(1,n1), rep(0,n0))
    X = rbind(data$s1, data$s0)
    colnames(X) = c(1:ncol(X))
    
    u_y_estimated <- test.surrogate(
      yone = data$y1, yzero = data$y0, sone = data$y1, szero = data$y0, epsilon = 0.1
    )$u.y
    # u_y_true = truth$uy_true
    # choose epsilon as maximum of estimated value of UY - 0.5 and 0
    eps <- max(0, u_y_estimated - 0.5)
    
    screen_res = rise_screen(Y, X, A, reference = "0", 
                             alpha = 0.05, epsilon = eps,
                             p_correction = "none", cores = 1)
    
    weights = data.frame("marker" = colnames(X),
                         "weight" = abs(1/screen_res$results[,"delta"]))
    res_evaluate = rise_validate(Y_validate = Y, X_validate = X, A_validate = A,
                                 markers = colnames(X), power_desired = 0.8,
                                 weights = weights, plot = F, alpha = 0.05)
    
    p_evaluate_df[j,i] = res_evaluate$validation_results[,"p_unadjusted"]
    print(paste0("Prop. invalid = ", prop_invalid_grid[i],  " : Simulation ", j, " of ", n_sim, " complete."))
  }
}
save(p_evaluate_df, file = "./output/simulation/metrics_figure4s.Rdata")
p_evaluate_df = get(load("./output/simulation/metrics_figure4s.Rdata"))

p_evaluate_long <- p_evaluate_df %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  ) %>%
  mutate(variable = as.numeric(variable))  # Convert variable names to numeric

# Create the violin plot
fig4 = ggplot(p_evaluate_long, aes(x = as.factor(variable), y = value, fill = as.factor(variable))) +
  geom_boxplot(outlier.shape = NA) +
  labs(
    x = "Proportion of invalid surrogates",
    y = "P-value",
    title = expression("Evaluation stage P-values as a function of the proportion of false positives in " * gamma[S])
  ) +
  geom_hline(yintercept = 0.05, linetype = "dashed", colour = "red", linewidth = 1.5) + 
  theme_minimal(base_size = 26) +
  theme(plot.title = element_blank(),
        plot.background = element_rect(fill = 'white', color = 'white'),
        legend.position = "none") +
  scale_fill_brewer(palette = "Spectral")

fig4


ggsave(plot = fig4, filename = "figure4.eps",
       path = "./output/figures/",
       units = "cm",
       width = 35,
       height = 20, dpi = 800)
```

# Supplementary Material
## Figure S1 
Distribution of p-values at the boundary value under the null.

```{r}
alpha <- 0.05 # nominal significance level
beta <- 0.2 # desired power for univariate surrogate test = (1-beta)*100% 
s0_mean <- 0 # invalid surrogate mean value in untreated group
s0_sd <- 1 # invalid surrogate  sd value in untreated group
s1_mean <- 0 # invalid surrogate mean value in treated group
s1_sd <- 1 # invalid surrogate sd value in treated group
y0_mean <- 0 # primary response mean value in untreated group
y0_sd <- 1 # primary response sd value in untreated group
y1_mean <- 3 # primary response mean value in treated group
y1_sd <- 1 # primary response sd value in treated group
n <- 50 # total sample size
n0 <- n / 2 # treated sample size 
n1 <- n / 2 # untreated sample size 
p <- 500 # total number of predictors
prop_valid <- 0 # proportion of predictors which are valid surrogates
p_valid <- prop_valid * p # number of valid surrogates
p_invalid = (1 - prop_valid) * p # number of invalid surrogates
n_sim = 1000 # number of simulations
corr = 0 # correlation - off-diagonal element in surrogate covariance matrix

p_figs1 = matrix(0)

for (j in 1:n_sim){
    res <- simulate_results(n1 = n / 2, 
                            n0 = n / 2, 
                            p = p, 
                            prop_valid = 0, 
                            n_sim = 1, 
                            corr = corr, 
                            mode = "simple")
    # Append p-values to results 
    p_figs1 = append(p_figs1,res[["p_values"]][["p_unadjusted"]][1,] %>% as.matrix())
    
    print(paste0("simulation ", j, " of ", n_sim, " complete." ))
  }

metrics_figures1 =  data.frame("p" = p_figs1[-1])

save(p_figs1, file = "./output/simulation/metrics_figures1.Rdata")
metrics_figures1 = get(load("./output/simulation/metrics_figures1.Rdata"))

fig_s1 = metrics_figures1 %>% ggplot(aes(x = p)) +
  geom_histogram(binwidth = 0.03, fill = "lightblue", color = "black", aes(y = ..density..)) +
  labs(title = "Distribution of P-values under the null hypothesis", x = "P-value", y = "Density") +
  xlim(0,1) +
  theme_minimal(base_size = 20) +
  theme(plot.title = element_blank(),
        plot.background = element_rect(fill = 'white', color = 'white'))

fig_s1

ggsave(plot = fig_s1,
       "./output/figures/figureA1.eps", 
       units = "cm", 
       width = 25, 
       height = 7, dpi = 800)
```

## Figure S2 
Empirical power and proportion of false positives as a function of surrogate strength under different correction methods for a fixed sample size.
```{r}
# Set up the values to simulate
# Set parameters
alpha <- 0.05 # nominal significance level
beta <- 0.2 # desired power for univariate surrogate test = (1-beta)*100% 
s0_mean <- 0 # invalid surrogate mean value in untreated group
s0_sd <- 1 # invalid surrogate  sd value in untreated group
s1_mean <- 0 # invalid surrogate mean value in treated group
s1_sd <- 1 # invalid surrogate sd value in treated group
y0_mean <- 0 # primary response mean value in untreated group
y0_sd <- 1 # primary response sd value in untreated group
y1_mean <- 3 # primary response mean value in treated group
y1_sd <- 1 # primary response sd value in treated group
n <- 50 # total sample size
n0 <- n / 2 # treated sample size 
n1 <- n / 2 # untreated sample size 
p = 500 # total number of predictors
prop_valid <- 0.1 # proportion of predictors which are valid surrogates
p_valid <- prop_valid * p # number of valid surrogates
p_invalid = (1 - prop_valid) * p # number of invalid surrogates
n_sim = 500 # number of simulations
corr = 0 # correlation - off-diagonal element in surrogate covariance matrix
sigma_grid = c(0.01,0.65,1.8,3,5.5,9,15,30,68,244) # values of sigma

# Initialise an empty list for p-values and dataframe for metrics
p_values = list()
metrics_figs2 <- data.frame(
  n = integer(), 
  p = integer(), 
  sigma_S = numeric(),
  avg_us = numeric(),
  avg_fpr_unadjusted = numeric(),
  avg_fdr_unadjusted = numeric(),
  avg_tpr_unadjusted = numeric(),
  avg_ppv_unadjusted = numeric(),
  avg_fpr_bonf = numeric(),
  avg_fdr_bonf = numeric(),
  avg_tpr_bonf = numeric(),
  avg_ppv_bonf = numeric(),
  avg_fpr_bh = numeric(),
  avg_fdr_bh = numeric(),
  avg_tpr_bh = numeric(),
  avg_ppv_bh = numeric(),
  avg_fpr_by = numeric(),
  avg_fdr_by = numeric(),
  avg_tpr_by = numeric(),
  avg_ppv_by = numeric()
)

# Perform simulations over grid
for (sigma in sigma_grid){
  avg_us = round(mean(calc.truth(p, prop_valid, valid_sigma = sigma, corr, mode = "simple")$us_true),2)
  res <- simulate_results(n1 = n / 2, 
                          n0 = n / 2, 
                          p = p, 
                          prop_valid, 
                          n_sim = n_sim, 
                          valid_sigma = sigma, 
                          corr = corr, 
                          mode = "simple")
      
      # p_values = append(p_values,list(res[["p_values"]][["p_unadjusted"]]))
      
      # Append results to data frame
      metrics_figs2 <- rbind(metrics_figs2, data.frame(
        n = n, p = p,
        sigma_S = sigma,
        avg_us = avg_us,
        avg_fpr_unadjusted = res[["metrics"]][["metrics_unadjusted"]][["avg_fpr"]], 
        avg_fdr_unadjusted = res[["metrics"]][["metrics_unadjusted"]][["avg_fdr"]], 
        avg_tpr_unadjusted = res[["metrics"]][["metrics_unadjusted"]][["avg_tpr"]],
        avg_ppv_unadjusted = res[["metrics"]][["metrics_unadjusted"]][["avg_ppv"]],
        avg_fpr_bonf = res[["metrics"]][["metrics_bonf"]][["avg_fpr"]], 
        avg_fdr_bonf = res[["metrics"]][["metrics_bonf"]][["avg_fdr"]], 
        avg_tpr_bonf = res[["metrics"]][["metrics_bonf"]][["avg_tpr"]],
        avg_ppv_bonf = res[["metrics"]][["metrics_bonf"]][["avg_ppv"]],
        avg_fpr_bh = res[["metrics"]][["metrics_bh"]][["avg_fpr"]], 
        avg_fdr_bh = res[["metrics"]][["metrics_bh"]][["avg_fdr"]], 
        avg_tpr_bh = res[["metrics"]][["metrics_bh"]][["avg_tpr"]],
        avg_ppv_bh = res[["metrics"]][["metrics_bh"]][["avg_ppv"]],
        avg_fpr_by = res[["metrics"]][["metrics_by"]][["avg_fpr"]], 
        avg_fdr_by = res[["metrics"]][["metrics_by"]][["avg_fdr"]], 
        avg_tpr_by = res[["metrics"]][["metrics_by"]][["avg_tpr"]],
        avg_ppv_by = res[["metrics"]][["metrics_by"]][["avg_ppv"]]
      ))
      print(paste0("Simulation ", which(sigma_grid == sigma), " of ", length(sigma_grid), " complete."))
}

save(metrics_figs2, file = "./output/simulation/metrics_figures2.Rdata")

metrics_figs2 = get(load("./output/simulation/metrics_figures2.Rdata"))

# Pivot metrics to long format 
metrics_figs2_fdr_long <- metrics_figs2  %>%
  pivot_longer(cols = c(avg_fdr_unadjusted, 
                        avg_fdr_bonf, 
                        avg_fdr_bh, 
                        avg_fdr_by), 
               names_to = "Correction", 
               values_to = "Value") %>% 
  dplyr::select(avg_us, Correction, Value)

metrics_figs2_tpr_long <- metrics_figs2  %>%
  pivot_longer(cols = c(avg_tpr_unadjusted, 
                        avg_tpr_bonf, 
                        avg_tpr_bh, 
                        avg_tpr_by), 
               names_to = "Correction", 
               values_to = "Value") %>% 
  dplyr::select(avg_us, Correction, Value)

# Plot
FDR_figs2 = ggplot(metrics_figs2_fdr_long, aes(x = avg_us, y = Value, linetype = Correction, color = Correction)) +
  geom_line(size = 1.3) +
  scale_color_brewer(palette = "Set1") +
  labs(x = expression(bar(U[S])), y = "False Discovery Proportion", linetype = "Correction") +
  scale_linetype_manual(values = c("avg_fdr_unadjusted" = "solid", 
                                   "avg_fdr_bonf" = "dashed",
                                   "avg_fdr_bh" = "dotted",
                                   "avg_fdr_by" = "dotdash"),
                        labels = c("avg_fdr_unadjusted" = "Unadjusted", 
                                   "avg_fdr_bonf" = "Bonferroni",
                                   "avg_fdr_bh" = "B-H",
                                   "avg_fdr_by" = "B-Y")) +
  scale_color_manual(values = c("avg_fdr_unadjusted" = brewer.pal(n = 4, name = "Set1")[1], 
                                   "avg_fdr_bonf" = brewer.pal(n = 4, name = "Set1")[2],
                                   "avg_fdr_bh" = brewer.pal(n = 4, name = "Set1")[3],
                                   "avg_fdr_by" = brewer.pal(n = 4, name = "Set1")[4]),
                     labels = c("avg_fdr_unadjusted" = "Unadjusted", 
                                   "avg_fdr_bonf" = "Bonferroni",
                                   "avg_fdr_bh" = "B-H",
                                   "avg_fdr_by" = "B-Y")) + 
  theme_minimal(base_size = 24) +
  theme(plot.title = element_blank(),
        plot.background = element_rect(fill = 'white', color = 'white'))

TPR_figs2 = ggplot(metrics_figs2_tpr_long, aes(x = avg_us, y = Value, linetype = Correction, color = Correction)) +
  geom_line(size = 1.3) +
  scale_color_brewer(palette = "Set1") +
  labs(x = expression(bar(U[S])), y = "Empirical Power", linetype = "Correction") +
  scale_linetype_manual(values = c("avg_tpr_unadjusted" = "solid", 
                                   "avg_tpr_bonf" = "dashed",
                                   "avg_tpr_bh" = "dotted",
                                   "avg_tpr_by" = "dotdash"),
                        labels = c("avg_tpr_unadjusted" = "Unadjusted", 
                                   "avg_tpr_bonf" = "Bonferroni",
                                   "avg_tpr_bh" = "B-H",
                                   "avg_tpr_by" = "B-Y")) +
  scale_color_manual(values = c("avg_tpr_unadjusted" = brewer.pal(n = 4, name = "Set1")[1], 
                                   "avg_tpr_bonf" = brewer.pal(n = 4, name = "Set1")[2],
                                   "avg_tpr_bh" = brewer.pal(n = 4, name = "Set1")[3],
                                   "avg_tpr_by" = brewer.pal(n = 4, name = "Set1")[4]),
                     labels = c("avg_tpr_unadjusted" = "Unadjusted", 
                                   "avg_tpr_bonf" = "Bonferroni",
                                   "avg_tpr_bh" = "B-H",
                                   "avg_tpr_by" = "B-Y")) + 
  theme_minimal(base_size = 24) +
  theme(plot.title = element_blank(),
        plot.background = element_rect(fill = 'white', color = 'white'))

## FPR and TPR in the same grid
TPR_patchwork_figs2 = TPR_figs2  + 
  ylim(0,1) + 
  theme(legend.position = "none",
        plot.title = element_blank())

FDR_patchwork_figs2 = FDR_figs2  + 
  ylim(0,1) + 
  theme(legend.position = "none",
        plot.title = element_blank())

legend = get_legend(ggplot(metrics_figs2_tpr_long, aes(x = avg_us, y = Value, linetype = Correction, color = Correction)) +
  geom_line(size = 1.3) +
  scale_color_brewer(palette = "Set1") +
  labs(x = expression(bar(U[S])), y = "False Discovery Proportion", linetype = "Correction") +
  scale_linetype_manual(values = c("avg_tpr_unadjusted" = "solid", 
                                   "avg_tpr_bonf" = "dashed",
                                   "avg_tpr_bh" = "dotted",
                                   "avg_tpr_by" = "dotdash"),
                        labels = c("avg_tpr_unadjusted" = "Unadjusted", 
                                   "avg_tpr_bonf" = "Bonferroni",
                                   "avg_tpr_bh" = "B-H",
                                   "avg_tpr_by" = "B-Y")) +
  scale_color_manual(values = c("avg_tpr_unadjusted" = brewer.pal(n = 4, name = "Set1")[1], 
                                   "avg_tpr_bonf" = brewer.pal(n = 4, name = "Set1")[2],
                                   "avg_tpr_bh" = brewer.pal(n = 4, name = "Set1")[3],
                                   "avg_tpr_by" = brewer.pal(n = 4, name = "Set1")[4]),
                     labels = c("avg_tpr_unadjusted" = "Unadjusted", 
                                   "avg_tpr_bonf" = "Bonferroni",
                                   "avg_tpr_bh" = "B-H",
                                   "avg_tpr_by" = "B-Y")) + 
  theme_minimal(base_size = 20) +
  theme(plot.background = element_rect(fill = 'white', color = 'white'),
        legend.key.width = unit(2, "cm")))

figures2 = grid.arrange(
  arrangeGrob(TPR_patchwork_figs2, FDR_patchwork_figs2, legend, 
              ncol = 3, widths = c(1, 1, 0.3)))  # Adjust fontsize and fontface


figures2

ggsave(plot = figures2,
       "./output/figures/figureA2.eps", 
       units = "cm", 
       width = 41, 
       height = 18, dpi = 800)
```

## Figure S3
FDP and empirical power as a function of inter-predictor correlation for a fixed sample size and surrogate strength
```{r}
alpha <- 0.05 # nominal significance level
beta <- 0.2 # desired power for univariate surrogate test = (1-beta)*100% 
s0_mean <- 0 # invalid surrogate mean value in untreated group
s0_sd <- 1 # invalid surrogate sd value in untreated group
s1_mean <- 0 # invalid surrogate mean value in treated group
s1_sd <- 2 # invalid surrogate sd value in treated group
y0_mean <- 0 # primary response mean value in untreated group
y0_sd <- 1 # primary response sd value in untreated group
y1_mean <- 3 # primary response mean value in treated group
y1_sd <- 1 # primary response sd value in treated group
valid_sigma <- 1.8 # set sigma such that average surrogate strength = 0.9
n <- 50 # fixed total sample size
n0 <- n / 2 # treated sample size 
n1 <- n / 2 # untreated sample size 
p = 500 # total number of predictors
prop_valid <- 0.1 # proportion of predictors which are valid surrogates
p_valid <- prop_valid * p # number of valid surrogates
p_invalid = (1 - prop_valid) * p # number of invalid surrogates
n_sim = 500 # number of simulations

corr_grid = seq(0,0.5,0.1) # grid of correlation values
metrics_figs3 = matrix(0,nrow = length(corr_grid)*n_sim, ncol = 3)

for (i in 1:length(corr_grid)){
  for (j in 1:n_sim){
    res <- simulate_results(n1 = n / 2, 
                            n0 = n / 2, 
                            p = p, 
                            prop_valid, 
                            n_sim = 1, 
                            valid_sigma,
                            corr = corr_grid[i],
                            mode = "simple")
    
    # Append results to data frame
    
    metrics_figs3[(i-1)*(n_sim) + j,1] = corr_grid[i]
    metrics_figs3[(i-1)*(n_sim) + j,2] = res[["metrics"]][["metrics_unadjusted"]][["avg_fdr"]]
    metrics_figs3[(i-1)*(n_sim) + j,3] = res[["metrics"]][["metrics_unadjusted"]][["avg_tpr"]]
    
    
    print(paste0("Correlation : ", corr_grid[i], ", simulation ", j, " of ", n_sim, " complete." ))
  }
}

save(metrics_figs3, file = "./output/simulation/metrics_figures3.Rdata")
metrics_figs3 = get(load("./output/simulation/metrics_figures3.Rdata"))

metrics_figs3 = data.frame("correlation" = metrics_figs3[,1], "fdr" = metrics_figs3[,2], "tpr" = metrics_figs3[,3])

fdr_plot_s3 = metrics_figs3 %>% ggplot(aes(x = as.factor(correlation), y = fdr, fill = as.factor(correlation))) + 
                           geom_violin() +
                           ylim(0,1) + 
                           ylab("False Discovery Proportion") +
                           xlab(expression(sigma[corr])) + 
                           ggtitle("Proportion of False Positives as a function of inter-predictor correlation") +
                           theme_minimal(base_size = 24) +
                           theme(plot.background = element_rect(fill = 'white', color = 'white'),
                                 plot.title = element_blank(),
                                 legend.position = "none",
        axis.title.x = element_text(size = 35)) +
  scale_fill_brewer(palette = "RdYlGn")

tpr_plot_s3 = metrics_figs3 %>% ggplot(aes(x = as.factor(correlation), y = tpr, fill = as.factor(correlation))) + 
  geom_violin() +
  ylim(0,1) + 
  ylab("Empirical Power") +
  xlab(expression(sigma[corr])) + 
  ggtitle("Proportion of False Positives as a function of inter-predictor correlation") +
  theme_minimal(base_size = 24) +
  theme(plot.background = element_rect(fill = 'white', color = 'white'),
        plot.title = element_blank(),
        legend.position = "none",
        axis.title.x = element_text(size = 35)) +
  scale_fill_brewer(palette = "RdYlGn")

figures3 = grid.arrange(
  arrangeGrob(tpr_plot_s3, fdr_plot_s3,  
              ncol = 2, widths = c(1, 1)))

ggsave(plot = figures3,
       "./output/figures/figureA3.eps", 
       units = "cm", 
       width = 35, 
       height = 18, dpi = 800)
```

## Figure S4
Boxplots of FPR as a function of sample size in the uncorrelated setting, more complex data generation.

```{r}
# Set parameters
alpha <- 0.05 # nominal significance level
beta <- 0.2 # desired power for univariate surrogate test = (1-beta)*100% 
s0_mean <- 0 # invalid surrogate mean value in untreated group
s0_sd <- 1 # invalid surrogate sd value in untreated group
s1_mean <- 0 # invalid surrogate mean value in treated group
s1_sd <- 2 # invalid surrogate sd value in treated group
y0_mean <- 0 # primary response mean value in untreated group
y0_sd <- 1 # primary response sd value in untreated group
y1_mean <- 3 # primary response mean value in treated group
y1_sd <- 1 # primary response sd value in treated group
p <- 500 # total number of predictors
prop_valid <- 0 # proportion of predictors which are valid surrogates
p_valid <- prop_valid * p # number of valid surrogates
p_invalid = (1 - prop_valid) * p # number of invalid surrogates
n_sim = 500 # number of simulations
corr = 0

n_grid = seq(10,100,10)

metrics_figs4 = matrix(0,nrow = length(n_grid)*n_sim, ncol = 2)

for (i in 1:length(n_grid)){
  for (j in 1:n_sim){
    res <- simulate_results(n1 = n_grid[i] / 2, 
                            n0 = n_grid[i] / 2, 
                            p = p, 
                            prop_valid = 0, 
                            n_sim = 1, 
                            valid_sigma,
                            corr = 0, 
                            mode = "complex")
    
    metrics_figs4[(i-1)*(n_sim) + j,1] = n_grid[i]
    metrics_figs4[(i-1)*(n_sim) + j,2] = res[["metrics"]][["metrics_unadjusted"]][["avg_fpr"]]
    print(paste0("Sample size : ", n_grid[i], ", simulation ", j, " of ", n_sim, " complete." ))
  }
}

metrics_figs4 = data.frame("n" = metrics_figs4[,1],
                          "False Positive Rate" = metrics_figs4[,2])

save(metrics_figs4, file = "./output/simulation/metrics_figures4.Rdata")
metrics_figs4 = get(load("./output/simulation/metrics_figures4.Rdata"))
figs4 = metrics_figs4 %>% ggplot(aes(x = as.factor(n), 
                                   y = FPR, 
                                   fill = as.factor(n))) +
  geom_boxplot() +
  labs(title = "Observed FPR across different sample sizes - DGF 2",
       x = "Sample size",
       y = "False Positive Rate",
       fill = "Sample Size") + 
  ylim(0,0.15) +
  geom_hline(yintercept = 0.05, colour = "red", linetype = "dashed", linewidth = 1.1) +
  theme_minimal(base_size = 35) +
  theme(plot.title = element_blank(),
        plot.background = element_rect(fill = 'white', color = 'white'),
        legend.position = "none") +
  scale_fill_brewer(palette = "RdYlGn")

figs4

ggsave(plot = figs4,
       "./output/figures/figureA4.eps", 
       units = "cm", 
       width = 40, 
       height = 25, dpi = 800)
```

## Figure S5
Violin plots of FPR as a function of correlation, complex data generation.
```{r}
# Set parameters
alpha <- 0.05 # nominal significance level
beta <- 0.2 # desired power for univariate surrogate test = (1-beta)*100% 
s0_mean <- 0 # invalid surrogate mean value in untreated group
s0_sd <- 1 # invalid surrogate sd value in untreated group
s1_mean <- 0 # invalid surrogate mean value in treated group
s1_sd <- 2 # invalid surrogate sd value in treated group
y0_mean <- 0 # primary response mean value in untreated group
y0_sd <- 1 # primary response sd value in untreated group
y1_mean <- 3 # primary response mean value in treated group
y1_sd <- 1 # primary response sd value in treated group
n <- 50 # total sample size
n0 <- n / 2 # treated sample size 
n1 <- n / 2 # untreated sample size 
p = 500 # total number of predictors
prop_valid <- 0 # proportion of predictors which are valid surrogates
p_valid <- prop_valid * p # number of valid surrogates
p_invalid = (1 - prop_valid) * p # number of invalid surrogates
n_sim = 500 # number of simulations

corr_grid = seq(0,0.5,0.1)

metrics_figs5 = matrix(0,nrow = length(corr_grid)*n_sim, ncol = 2)

for (i in 1:length(corr_grid)){
  for (j in 1:n_sim){
      res <- simulate_results(n1 = n / 2, 
                              n0 = n / 2, 
                              p = p, 
                              prop_valid = 0, 
                              n_sim = 1, 
                              corr = corr_grid[i], 
                              mode = "complex")
      
      # p_values = append(p_values,list(res[["p_values"]][["p_unadjusted"]]))
      
      # Append results to data frame
      
      metrics_figs5[(i-1)*(n_sim) + j,1] = corr_grid[i]
      metrics_figs5[(i-1)*(n_sim) + j,2] = res[["metrics"]][["metrics_unadjusted"]][["avg_fpr"]]
      print(paste0("Correlation : ", corr_grid[i], ", simulation ", j, " of ", n_sim, " complete." ))
      }
}

metrics_figs5 = data.frame("correlation" = metrics_figs5[,1],
                          "False Positive Rate" = metrics_figs5[,2])

save(metrics_figs5, file = "./output/simulation/metrics_figures5.Rdata")

metrics_figs5 = get(load("./output/simulation/metrics_figures5.Rdata"))

figs5 = metrics_figs5 %>% ggplot(aes(x = as.factor(correlation), 
                                   y = FPR,
                                   fill = as.factor(correlation))) +
  geom_violin() +
  labs(title = "Observed FPR across different inter-predictor correlation levels - DGF 2",
       x = expression(sigma[corr]),
       y = "False Positive Rate",
       fill = "Correlation") + 
  geom_hline(yintercept = 0.05, colour = "red", linetype = "dashed", linewidth = 1.1) +
  theme_minimal(base_size = 35) +
  theme(plot.title = element_blank(),
        plot.background = element_rect(fill = 'white', color = 'white'),
        legend.position = "none",
        axis.title.x = element_text(size = 55)) +
  scale_fill_brewer(palette = "RdYlGn")

figs5

ggsave(plot = figs5,
       "./output/figures/figureA5.eps", 
       units = "cm", 
       width = 40, 
       height = 20, dpi = 800)
```

## Figure S6
Empirical power and proportion of false positives as a function of average surrogate strength for three different sample sizes, complex data generation. Average surrogate strengths corresponding to (0.55,0.6,...,0.95, 0.985) are found by trial-and-error.
```{r}
# round(mean(calc.truth(p = 500, prop_valid = 0.1, valid_sigma = 40000, corr, mode = "complex")$us_true),2) # 0.55
# round(mean(calc.truth(p = 500, prop_valid = 0.1, valid_sigma = 9000, corr, mode = "complex")$us_true),2) # 0.6
# round(mean(calc.truth(p = 500, prop_valid = 0.1, valid_sigma = 4000, corr, mode = "complex")$us_true),2) # 0.65
# round(mean(calc.truth(p = 500, prop_valid = 0.1, valid_sigma = 1600, corr, mode = "complex")$us_true),2) # 0.7
# round(mean(calc.truth(p = 500, prop_valid = 0.1, valid_sigma = 850, corr, mode = "complex")$us_true),2) # 0.75
# round(mean(calc.truth(p = 500, prop_valid = 0.1, valid_sigma = 400, corr, mode = "complex")$us_true),2) # 0.8
# round(mean(calc.truth(p = 500, prop_valid = 0.1, valid_sigma = 210, corr, mode = "complex")$us_true),2) # 0.85
# round(mean(calc.truth(p = 500, prop_valid = 0.1, valid_sigma = 80, corr, mode = "complex")$us_true),2) # 0.9
# round(mean(calc.truth(p = 500, prop_valid = 0.1, valid_sigma = 15, corr, mode = "complex")$us_true),2) # 0.95
# round(mean(calc.truth(p = 500, prop_valid = 0.1, valid_sigma = 0.0001, corr, mode = "complex")$us_true),3) # 0.985

# Set up the values to simulate
alpha <- 0.05 # nominal significance level
beta <- 0.2 # desired power for univariate surrogate test = (1-beta)*100% 
s0_mean <- 0 # invalid surrogate mean value in untreated group
s0_sd <- 1 # invalid surrogate sd value in untreated group
s1_mean <- 0 # invalid surrogate mean value in treated group
s1_sd <- 2 # invalid surrogate sd value in treated group
y0_mean <- 0 # primary response mean value in untreated group
y0_sd <- 1 # primary response sd value in untreated group
y1_mean <- 3 # primary response mean value in treated group
y1_sd <- 1 # primary response sd value in treated group
p = 500 # total number of predictors
sigma_grid = c(0.0001,15,80,210,400,850,1600,4000,9000,40000) # values of sigma
n_grid <- c(30,50,100) # 3 sample sizes
n_sim = 500
prop_valid = 0.1
corr = 0

# Initialise an empty list for p-values and dataframe for metrics
p_values = list()
metrics_figs6 <- data.frame(
  n = integer(), 
  p = integer(), 
  sigma_S = numeric(),
  avg_us = numeric(),
  avg_fpr_unadjusted = numeric(),
  avg_fdr_unadjusted = numeric(),
  avg_tpr_unadjusted = numeric(),
  avg_ppv_unadjusted = numeric(),
  avg_fpr_bonf = numeric(),
  avg_fdr_bonf = numeric(),
  avg_tpr_bonf = numeric(),
  avg_ppv_bonf = numeric(),
  avg_fpr_bh = numeric(),
  avg_fdr_bh = numeric(),
  avg_tpr_bh = numeric(),
  avg_ppv_bh = numeric(),
  avg_fpr_by = numeric(),
  avg_fdr_by = numeric(),
  avg_tpr_by = numeric(),
  avg_ppv_by = numeric()
)

# Perform simulations over grid
for (sigma in sigma_grid){
  avg_us = round(mean(calc.truth(p, prop_valid, valid_sigma = sigma, corr, mode = "complex")$us_true),2)
  for (n in n_grid) {
      res <- simulate_results(n1 = n / 2, 
                              n0 = n / 2, 
                              p = p, 
                              prop_valid, 
                              n_sim = n_sim, 
                              valid_sigma = sigma,
                              corr = 0, 
                              mode = "complex")
      
      # p_values = append(p_values,list(res[["p_values"]][["p_unadjusted"]]))
      
      # Append results to data frame
      metrics_figs6 <- rbind(metrics_figs6, data.frame(
        n = n, p = p,
        sigma_S = sigma,
        avg_us = avg_us,
        avg_fpr_unadjusted = res[["metrics"]][["metrics_unadjusted"]][["avg_fpr"]], 
        avg_fdr_unadjusted = res[["metrics"]][["metrics_unadjusted"]][["avg_fdr"]], 
        avg_tpr_unadjusted = res[["metrics"]][["metrics_unadjusted"]][["avg_tpr"]],
        avg_ppv_unadjusted = res[["metrics"]][["metrics_unadjusted"]][["avg_ppv"]],
        avg_fpr_bonf = res[["metrics"]][["metrics_bonf"]][["avg_fpr"]], 
        avg_fdr_bonf = res[["metrics"]][["metrics_bonf"]][["avg_fdr"]], 
        avg_tpr_bonf = res[["metrics"]][["metrics_bonf"]][["avg_tpr"]],
        avg_ppv_bonf = res[["metrics"]][["metrics_bonf"]][["avg_ppv"]],
        avg_fpr_bh = res[["metrics"]][["metrics_bh"]][["avg_fpr"]], 
        avg_fdr_bh = res[["metrics"]][["metrics_bh"]][["avg_fdr"]], 
        avg_tpr_bh = res[["metrics"]][["metrics_bh"]][["avg_tpr"]],
        avg_ppv_bh = res[["metrics"]][["metrics_bh"]][["avg_ppv"]],
        avg_fpr_by = res[["metrics"]][["metrics_by"]][["avg_fpr"]], 
        avg_fdr_by = res[["metrics"]][["metrics_by"]][["avg_fdr"]], 
        avg_tpr_by = res[["metrics"]][["metrics_by"]][["avg_tpr"]],
        avg_ppv_by = res[["metrics"]][["metrics_by"]][["avg_ppv"]]
      ))
    print(paste0("Sigma = ", sigma, ", n = ", n, " complete."))
  }
}

save(metrics_figs6, file = "./output/simulation/metrics_figures6.Rdata")

metrics_figs6 = get(load("./output/simulation/metrics_figures6.Rdata"))

# FDR as a function of surrogate strength
FDR_plot_unadjusted = metrics_figs6 %>% ggplot(aes(x = avg_us, y = avg_fdr_unadjusted, color = as.factor(n), linetype = as.factor(n))) + 
  geom_line(size = 1.3) + 
  geom_point() +
  ylab("False Discovery Proportion") +
  xlab(expression(bar(U[S]))) +
  scale_color_brewer(palette = "Set1") + 
  guides(color = guide_legend(title = "Sample size")) +
  ggtitle("Proportion of False Positives as a function of average surrogate strength - unadjusted p-values") +
  theme_minimal(base_size = 24) +
  theme(plot.background = element_rect(fill = 'white', color = 'white'),
        plot.title = element_blank()) 

TPR_plot_unadjusted  = metrics_figs6 %>% ggplot(aes(x = avg_us, y = avg_tpr_unadjusted , color = as.factor(n), linetype = as.factor(n))) + 
  geom_line(size = 1.3) + 
  geom_point() +
  ylab("Empirical power") +
  xlab(expression(bar(U[S]))) +
  scale_color_brewer(palette = "Set1") + 
  guides(color = guide_legend(title = "Sample size")) +
  ggtitle("Empirical power as a function of average surrogate strength - unadjusted p-values") +
  theme_minimal(base_size = 24) +
  theme(plot.background = element_rect(fill = 'white', color = 'white'),
        plot.title = element_blank()) 

## FPR and TPR in the same grid 
TPR_patchwork = TPR_plot_unadjusted + 
  ylim(0,1) + 
  theme(legend.position = "none")

FDR_patchwork = FDR_plot_unadjusted + 
  ylim(0,1) + 
  theme(legend.position = "none")

# Extract the legend
legend <- get_legend(
  metrics_figs6 %>% 
    ggplot(aes(x = avg_us, 
               y = avg_fdr_unadjusted, 
               color = as.factor(n), 
               linetype = as.factor(n))) + 
    geom_line(size = 1.3) + # Adjust the line size in the plot
    geom_point() +
    ylab("False Discovery Proportion") +
    xlab(expression(bar(U[S]))) +
    scale_color_brewer(palette = "Set1") + 
    guides(
      color = guide_legend(title = "Sample size"),
      linetype = guide_legend(
        title = "Sample size",
        override.aes = list(size = 1) # Adjust line size in legend
      )
    ) +
    ggtitle("Proportion of False Positives as a function of average surrogate strength - unadjusted p-values") +
    theme_minimal(base_size = 20) +
    theme(
      legend.key.width = unit(2, "cm") # Increase legend key width
    )
)

figures6 = grid.arrange(
  arrangeGrob(TPR_patchwork, FDR_patchwork, legend, 
              ncol = 3, widths = c(1, 1, 0.3)),  # Adjust widths as needed
  top = textGrob("",
                 gp = gpar(fontsize = 20))  # Adjust fontsize and fontface
)

ggsave(plot = figures6,
       "./output/figures/figureA6.eps", 
       units = "cm", 
       width = 35, 
       height = 18, dpi = 800)
```

## Figure S7 
Evaluation stage : Vary the proportion of false positives in the combined marker and examine distribution of p-values. Complex data generation. 

```{r}
# Step 1 : construct a marker as a combination of true and false positives
# Keep true positives strong on average : U_Y = 0.9
source("./R/rise_validate.R")
source("./R/rise_screen.R")
alpha <- 0.05 # nominal significance level
beta <- 0.2 # desired power for univariate surrogate test = (1-beta)*100% 
s0_mean <- 0 # invalid surrogate mean value in untreated group
s0_sd <- 1 # invalid surrogate sd value in untreated group
s1_mean <- 0 # invalid surrogate mean value in treated group
s1_sd <- 2 # invalid surrogate sd value in treated group
y0_mean <- 0 # primary response mean value in untreated group
y0_sd <- 1 # primary response sd value in untreated group
y1_mean <- 3 # primary response mean value in treated group
y1_sd <- 1 # primary response sd value in treated group
p <- 20 # total number of predictors
corr = 0 # inter-predictor correlation
n1 = 25 # treated individuals
n0 = 25 # untreated
p = 20 # number of genes to combine for the signature
valid_sigma = 80 # this corresponds to avg U_Y = 0.9
prop_invalid_grid = seq(0,1,0.1) # proportions of invalid surrogates in gamma to test
n_sim = 500 # number of simulations

# evaluate true surrogate strength
# truth <- calc.truth(p = 500, prop_valid=1, valid_sigma = 80, corr, mode = "complex")
# round(mean(truth$us_true),2)

p_evaluate_df = as.data.frame(matrix(0,nrow = n_sim, ncol = length(prop_invalid_grid)))
colnames(p_evaluate_df) = prop_invalid_grid
for (i in 1:length(prop_invalid_grid)){
  for (j in 1:n_sim){
    prop_invalid = prop_invalid_grid[i]
    prop_valid = 1 - prop_invalid 
    
    data = gen.data(n1, n0, p, prop_valid, valid_sigma, corr, mode = "complex")
    
    # get weights from screening function
    Y = c(data$y1,data$y0)
    A = c(rep(1,n1), rep(0,n0))
    X = rbind(data$s1, data$s0)
    colnames(X) = c(1:ncol(X))
    
    u_y_estimated <- test.surrogate(
      yone = data$y1, yzero = data$y0, sone = data$y1, szero = data$y0, epsilon = 0.1
    )$u.y
    # u_y_true = truth$uy_true
    # choose epsilon as maximum of estimated value of UY - 0.5 and 0
    eps <- max(0, u_y_estimated - 0.5)
    
    screen_res = rise_screen(Y, X, A, reference = "0", 
                             alpha = 0.05, epsilon = eps,
                             p_correction = "none", cores = 1)
    
    weights = data.frame("marker" = colnames(X),
                         "weight" = abs(1/screen_res$results[,"delta"]))
    res_evaluate = rise_validate(Y_validate = Y, X_validate = X, A_validate = A,
                                 markers = colnames(X), power_desired = 0.8,
                                 weights = weights, plot = F, alpha = 0.05)
    
    p_evaluate_df[j,i] = res_evaluate$validation_results[,"p_unadjusted"]
    print(paste0("Prop. invalid = ", prop_invalid_grid[i],  " : Simulation ", j, " of ", n_sim, " complete."))
  }
}
save(p_evaluate_df, file = "./output/simulation/metrics_figures7s.Rdata")
p_evaluate_df = get(load("./output/simulation/metrics_figures7s.Rdata"))

p_evaluate_long <- p_evaluate_df %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  ) %>%
  mutate(variable = as.numeric(variable))  # Convert variable names to numeric

# Create the violin plot
figs7 = ggplot(p_evaluate_long, aes(x = as.factor(variable), y = value, fill = as.factor(variable))) +
  geom_boxplot(outlier.shape = NA) +
  labs(
    x = "Proportion of invalid surrogates",
    y = "P-value",
    title = expression("Evaluation stage P-values as a function of the proportion of false positives in " * gamma[S])
  ) +
  geom_hline(yintercept = 0.05, linetype = "dashed", colour = "red", linewidth = 1.5) + 
  theme_minimal(base_size = 26) +
  theme(plot.title = element_blank(),
        plot.background = element_rect(fill = 'white', color = 'white'),
        legend.position = "none") +
  scale_fill_brewer(palette = "Spectral")

figs7


ggsave(plot = figs7, filename = "figureA7.eps",
       path = "./output/figures/",
       units = "cm",
       width = 35,
       height = 20, dpi = 800)
```

